---
title: "Banana - Striatum Lipids"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE,warnings=F,error = F)

library(ggplot2)
library(ggforce)
library(limma)
library(ggrepel)
library(imputeLCMD)
library(ggsci)
library(factoextra)
library(emmeans)
library(patchwork)
library(pheatmap)
library(brglm2)


```


# Section {.tabset}

## Read in data {.tabset}

```{r, include=F}

## Please select the path to the data file you want to analyze
data = read.csv("~/bbc-secondary/research/MOOD_yue_MS_template_VBCS1110/Plasma_metabolites/MC00279_merged_checked.csv",header=T)
colnames(data)[1] = "Sample"
# This line makes all of the compound variables numeric, sometimes they read in as characters
data[,2:ncol(data)] <- sapply( data[,2:ncol(data)], as.numeric )


if(any(grepl("blank",tolower(data$Sample),fixed=T))){
  # Now we'll do blank filtering
  # Start by making a data.frame of just the blank which we'll use for subtraction
  blank = log2(data[grepl("blank",tolower(data$Sample),fixed=T),-1])
  
  #Make all missing and -Inf values 0 for the purposes of easily checking for > 3 fold difference
  blank[which(apply(blank,2,function(x) sum(is.finite(x)))<1)] = 0
  
  #Substract actual valus from blank values, both on the log2 scale
  diff = as.data.frame(t(log2(t(data[,2:ncol(data)])) - as.vector(t(blank))))
  
  #Any difference that is less than 3 fold, make missing
  data[,2:ncol(data)][which(diff < log2(3),arr.ind=T)] = NA
}

#remove QC and process blanks
data = data[! (grepl("qc",tolower(data$Sample),fixed=T) |grepl("blank",tolower(data$Sample),fixed=T)) , ]

Molecule = colnames(data)[2:ncol(data)]

# Looking for Sex, genotypes; where is the sex information?
data$Sex =  sapply(strsplit(data$Sample,"_"), function(x) x[2])
data$Geno =  sapply(strsplit(data$Sample,"_"), function(x) x[3])
table(data$Geno,data$Sex)




```



### Missingness

Here we create a heatmap of missingness to see if it specific to any group of interest (Sex, Genotype). If there is any specificity, you would see very clear clustering blocks. If nothing shows up, there were no missing values


```{r, fig.height = 9,fig.width=9}
#Replace 0s with NA
data[data==0]=NA

miss.df = data
rownames(miss.df) = miss.df$Sample
miss.df = miss.df[,colnames(miss.df) %in% Molecule]
miss.df[!is.na(miss.df)] = 1
miss.df[is.na(miss.df)] = 0

missing = apply(miss.df,2, function(x) sum(x)/length(x))
anno = data.frame(Genotype = data$Geno,Sex = data$Sex)
rownames(anno) = rownames(miss.df)

if( length(which(missing<0.95)) > 0){
  pheatmap(t(miss.df[,which(missing<0.95)
  ]),color=c("white","black"),annotation_col = anno)
}

```


### Testing for specificity

```{r, include=F}

miss.df$Geno = data$Geno
miss.df$Sex = data$Sex
colnames(miss.df) = make.names(colnames(miss.df))
res.Sex=NULL
res.geno=NULL
res.pooled=NULL

for(i in colnames(miss.df[,which(missing<0.95)])){
  
  f=as.formula(paste0(i,"~ Sex*Geno"))
  fit = glm(f, data = miss.df, family = "binomial", method = brglmFit)
  
  res.Sex = rbind(res.Sex,data.frame(Compound = i,summary(emmeans(fit,pairwise~Sex|Geno,adjust="none")$contrasts)))
  
  res.geno = rbind(res.geno,data.frame(Compound = i,summary(emmeans(fit,pairwise~Geno|Sex,adjust="none")$contrasts)))
  
  res.pooled = rbind(res.geno,data.frame(Compound = i,summary(emmeans(fit,pairwise~Geno,adjust="none")$contrasts)))
}


res.geno$FDR = p.adjust(res.geno$p.value,method="BH")

res.Sex$FDR = p.adjust(res.Sex$p.value,method="BH")

res.pooled$FDR = p.adjust(res.pooled$p.value,method="BH")



```

#### If rows are empty or nothing is showing, no specificty was detected

```{r}

if(length(colnames(miss.df[,which(missing<0.95)]))>0){
  knitr::kable(res.Sex[res.Sex$FDR < 0.1,],caption="Compounds specific to a Sex")
  knitr::kable(res.geno[res.geno$FDR < 0.1,],caption="Compounds specific to a Genotype")
  knitr::kable(res.pooled[res.pooled$FDR < 0.1,],caption="Compounds specific to a Sex/Genotype Combo")
}


```

### Removed for too much missgingness

```{r}
#These compounds are being removed for excessive missingness
knitr::kable(names(missing[missing<0.3]))

#drop heavily missing data (>30%)
data = data[, ! colnames(data) %in% names(missing[missing<0.3])]

```


```{r}

#impute
data[,colnames(data) %in% Molecule] = t(imputeLCMD::impute.QRILC(log2(t(data[,colnames(data) %in% Molecule])))[[1]])


```

## Exploratory {.tabset}

### PCA

```{r, fig.height=6,fig.width=16}

pc.plot = function(df){

  pc = prcomp(data.matrix(df[,colnames(df) %in% Molecule]))

  pc.df = as.data.frame(pc$x)
  pc.df$Genotype = df$Geno
  pc.df$Sex = df$Sex

  p1 = ggplot(pc.df,aes(x=PC1,y=PC2,color = Genotype, linetype=Genotype )) +
    geom_point() +
    geom_mark_ellipse() +
    theme_bw(12) +
    scale_color_nejm()

  p2 = ggplot(pc.df,aes(x=PC3,y=PC4,color =Genotype, linetype=Genotype )) +
    geom_point() +
    geom_mark_ellipse() +
    theme_bw(12) +
    scale_color_nejm()

  p3 = ggplot(pc.df,aes(x=PC5,y=PC6,color=Genotype, linetype=Genotype )) +
    geom_point() +
    geom_mark_ellipse() +
    theme_bw(12) +
    scale_color_nejm()

  return(p1+p2+p3)

}

pc.plot(data) +  plot_annotation('Pooled Sexes',theme=theme(plot.title=element_text(hjust=0.5)))


pc.plot(data[data$Sex == "M",]) +  plot_annotation('Males Only',theme=theme(plot.title=element_text(hjust=0.5)))

pc.plot(data[data$Sex == "F",]) +  plot_annotation('Females Only',theme=theme(plot.title=element_text(hjust=0.5)))


```


### Heatmap {.tabset}

#### Pooled

```{r, fig.width=9,fig.height=25}

anno = data.frame(Genotype = data$Geno,Sex = data$Sex)
rownames(anno) = data$Sample
rownames(data) = data$Sample
pheatmap(t(log2(data[,colnames(data) %in% Molecule])),
         annotation_col = anno,
         scale="row",
         color=colorRampPalette(c("blue","white","red"))(100))


```


#### Males Only

```{r, fig.width=9,fig.height=25}

anno = data.frame(Genotype = data[data$Sex == "M",]$Geno,Sex = data[data$Sex == "M",]$Sex)
rownames(anno) = data[data$Sex == "M",]$Sample

pheatmap(t(log2(data[data$Sex == "M",colnames(data) %in% Molecule])),
         annotation_col = anno,
         scale="row",
         color=colorRampPalette(c("blue","white","red"))(100))



```



#### Females Only

```{r, fig.width=9,fig.height=25}

anno = data.frame(Genotype = data[data$Sex == "F",]$Geno,Sex = data[data$Sex == "F",]$Sex)
rownames(anno) = data[data$Sex == "F",]$Sample

pheatmap(t(log2(data[data$Sex == "F",colnames(data) %in% Molecule])),
         annotation_col = anno,
         scale="row",
         color=colorRampPalette(c("blue","white","red"))(100))



```


## Differntial Abundance {.tabset}
### Sex Stratified {.tabset}

```{r, fig.height=14,fig.width=14}


data$Geno_Sex = paste0(data$Geno,data$Sex)

design = model.matrix(~ 0 + Geno_Sex, data= data)


fit <- lmFit(t(data[,colnames(data) %in% Molecule]), design, robust = TRUE)

cont2 = makeContrasts( Geno_SexkikiF  - Geno_SexkiwtF ,
                       Geno_SexkikiF - Geno_SexwtwtF ,
                       Geno_SexkiwtF  - Geno_SexwtwtF,
                       
                       Geno_SexkikiM  - Geno_SexkiwtM ,
                       Geno_SexkikiM - Geno_SexwtwtM ,
                       Geno_SexkiwtM  - Geno_SexwtwtM,
                       
                       
                      levels = design)

set.seed(777)
fit.cont <- contrasts.fit(fit, contrasts = cont2) 

set.seed(777)
ebayesf <- eBayes(fit.cont, robust = T, trend = T)

tops.sig = topTable(ebayesf, adjust.method = "BH", p.value = 0.05, number = Inf)
tops.all = topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf)


```

#### Females


```{r, fig.height=6,fig.width=14}

f.df = rbind(data.frame(Contrast = "KI/KI - KI/WT", topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf,coef=1)),
               data.frame(Contrast = "KI/KI - WT/WT", topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf,coef=2)),
               data.frame(Contrast = "KI/WT - WT/WT", topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf,coef=3))
               )



f.df$Compound = rownames(f.df)
f.df$Significant = ifelse(f.df$adj.P.Val<0.05,"Significant","NS")
f.df$label = ifelse(f.df$Significant == "NS","",f.df$Compound)

ggplot(f.df,aes(x=logFC,y=-log10(P.Value),color=Significant)) + geom_point()+
  theme_bw(16) +facet_wrap(~Contrast)+geom_text_repel(data = f.df[f.df$label!= "",] ,aes(label=Compound))+scale_color_manual(values = c("NS"="black","Significant" = "red")) + ggtitle("HFD KI - WT")

knitr::kable(f.df[f.df$adj.P.Val<0.05,],row.names = F,digits = 4)


```

#### Males


```{r, fig.height=6,fig.width=14}

m.df = rbind(data.frame(Contrast = "KI/KI - KI/WT", topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf,coef=4)),
               data.frame(Contrast = "KI/KI - WT/WT", topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf,coef=5)),
               data.frame(Contrast = "KI/WT - WT/WT", topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf,coef=6))
               )



m.df$Compound = rownames(m.df)
m.df$Significant = ifelse(m.df$adj.P.Val<0.05,"Significant","NS")
m.df$label = ifelse(m.df$Significant == "NS","",m.df$Compound)

ggplot(m.df,aes(x=logFC,y=-log10(P.Value),color=Significant)) + geom_point()+
  theme_bw(16) +facet_wrap(~Contrast)+geom_text_repel(data = m.df[m.df$label!= "",] ,aes(label=Compound))+scale_color_manual(values = c("NS"="black","Significant" = "red")) + ggtitle("HFD KI - WT")

knitr::kable(m.df[m.df$adj.P.Val<0.05,],row.names = F,digits = 4)


```


#### Boxplots Top Compounds

```{r, fig.height=12,fig.width=16}

  plot.df = reshape2::melt(data[,colnames(data) %in% c("Geno","Sex",rownames(tops.all)[1:20])],id.vars=c("Geno","Sex"))
  title = "Boxplots of Compounds with largest differences"


ggplot( plot.df,aes(x = Geno,y=value,color=Sex)) +geom_boxplot(position = position_dodge()) +
  facet_wrap(~variable,scales="free_y")+theme_bw(16)+scale_color_aaas()+ggtitle(title)

```



### Pooled {.tabset}

#### Volcano


```{r, fig.height=6,fig.width=14}




data$Geno_Sex = paste0(data$Geno,data$Sex)

design = model.matrix(~ 0 + Geno+Sex, data= data)


fit <- lmFit(t(data[,colnames(data) %in% Molecule]), design, robust = TRUE)

cont2 = makeContrasts( Genokiki  - Genokiwt ,
                       Genokiki - Genowtwt ,
                       Genokiwt  - Genowtwt,
                       
                      levels = design)

set.seed(777)
fit.cont <- contrasts.fit(fit, contrasts = cont2) 

set.seed(777)
ebayesf <- eBayes(fit.cont, robust = T, trend = T)

tops.sig = topTable(ebayesf, adjust.method = "BH", p.value = 0.05, number = Inf)
tops.all = topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf)

m.df = rbind(data.frame(Contrast = "KI/KI - KI/WT", topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf,coef=1)),
               data.frame(Contrast = "KI/KI - WT/WT", topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf,coef=2)),
               data.frame(Contrast = "KI/WT - WT/WT", topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf,coef=3))
               )



m.df$Compound = rownames(m.df)
m.df$Significant = ifelse(m.df$adj.P.Val<0.05,"Significant","NS")
m.df$label = ifelse(m.df$Significant == "NS","",m.df$Compound)

ggplot(m.df,aes(x=logFC,y=-log10(P.Value),color=Significant)) + geom_point()+
  theme_bw(16) +facet_wrap(~Contrast)+geom_text_repel(data = m.df[m.df$label!= "",] ,aes(label=Compound))+scale_color_manual(values = c("NS"="black","Significant" = "red")) + ggtitle("HFD KI - WT")

knitr::kable(m.df[m.df$adj.P.Val<0.05,],row.names = F,digits = 4)




```


#### Boxplots Top Compounds

```{r, fig.height=12,fig.width=16}


  plot.df = reshape2::melt(data[,colnames(data) %in% c("Geno","Sex",rownames(tops.all)[1:20])],id.vars=c("Geno","Sex"))
  title = "Boxplots of Compounds with largest differences"



ggplot( plot.df,aes(x = Geno,y=value)) +geom_boxplot(position = position_dodge()) +
  facet_wrap(~variable,scales="free_y")+theme_bw(16)+scale_color_aaas()+ggtitle(title)

```

