---
title: "HFD - Striatum Lipids"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE,warnings=F,error = F)

library(ggplot2)
library(ggforce)
library(limma)
library(dplyr)
library(ggrepel)
library(imputeLCMD)
library(ggsci)
library(factoextra)
library(emmeans)
library(patchwork)
library(pheatmap)
library(brglm2)
library(readxl)

#Convenience function for plotting the volcano plot and saving the results
plot_table_save = function(data){
  data$Compound = rownames(data)
  data$Significant = ifelse(data$adj.P.Val<0.05,"Significant","NS")
  data$label = ifelse(data$Significant == "NS","",data$Compound)
  data$Direction = ifelse(data$logFC>0,"Up regulated","Down regulated")
  data$Direction[data$Significant=="NS"] = "NS"
  
  volc = ggplot(data,aes(x=logFC,y=-log10(adj.P.Val),color=Direction)) + geom_point()+
  theme_bw(16) +facet_wrap(~Contrast)+geom_text_repel(data = data[data$label!= "",] ,aes(label=Compound))+scale_color_manual(values = c("Up regulated"="red","Down regulated" = "green","NS = Grey")) + ggtitle(data$Contrast)

  print(volc)
  
  ggsave(plot = volc,path=paste0(gsub("\\..*","", file)),filename = paste0(data$Contrast[1],"_volcano.png"),create.dir = T)
  print("\n")

knitr::kable(data[data$adj.P.Val<1,],row.names = F,digits = 4)

  write.csv(data,file =paste0(gsub("\\..*","", file),"/",data$Contrast[1],"_results.csv") )  
  
}

#Plots top 20 boxplots
plot_differences = function(df,level="Pooled"){
  plot.df = reshape2::melt(df[,colnames(df) %in% c("Geno","Diet",rownames(tops.all)[1:20])],id.vars=c("Geno","Diet"))
  title = "Boxplots of Compounds with largest differences"

  print(ggplot( plot.df,aes(x = Geno,y=value,color=Diet)) +geom_boxplot(position = position_dodge()) +
    facet_wrap(~variable,scales="free_y")+theme_bw(16)+scale_color_aaas()+ggtitle(title))

  ggsave(path=paste0(gsub("\\..*","", file)),filename = paste0(level,"_top_boxplot.png"),create.dir = T)
  
}
 


```


# Section {.tabset}

## Read in data {.tabset}

```{r, include=F}

file = "~/bbc-secondary/research/MOOD_yue_MS_template_VBCS1110/20240118_MC00470_profiling_KidneyLipids/3_tables/MC00470_Merged_PoolSize.csv"

## Please select the path to the data file you want to analyze
data = read.csv(file,header=T)


## Update the file path for the Excel file with Sex info, this will need to  be merged into these files since the MS files don't seem to have sex info
meta = subset(readxl::read_excel("~/bbc-secondary/research/MOOD_yue_MS_template_VBCS1110/sample submittion_20231030_Banana_HFD aged.xlsx",sheet=3,skip=1),select=c('Core ID','Condition 3'))
colnames(meta) = c("ID","Sex")
meta$ID = paste0("S",meta$ID)
meta$Sex = toupper(meta$Sex)
colnames(data)[1] = "Sample"

# This line makes all of the compound variables numeric, sometimes they read in as characters
data[,2:ncol(data)] <- sapply( data[,2:ncol(data)], as.numeric)

if(any(grepl("blank",tolower(data$Sample),fixed=T))){
  # Now we'll do blank filtering
  # Start by making a data.frame of just the blank which we'll use for subtraction
  blank = log2(data[grepl("blank",tolower(data$Sample),fixed=T),-1])
  
  #Make all missing and -Inf values 0 for the purposes of easily checking for > 3 fold difference
  blank[which(apply(blank,2,function(x) sum(is.finite(x)))<1)] = 0
  
  #Subtract actual values from blank values, both on the log2 scale
  diff = as.data.frame(t(log2(t(data[,2:ncol(data)])) - as.vector(t(blank))))
  
  #Any difference that is less than 3 fold, make missing
  data[,2:ncol(data)][which(diff < log2(3),arr.ind=T)] = NA
}

#remove QC and process blanks
data = data[! (grepl("qc",tolower(data$Sample),fixed=T) |grepl("blank",tolower(data$Sample),fixed=T)) , ]

Molecule = colnames(data)[2:ncol(data)]

# Looking for diet, genotypes; where is the sex information?
data$Geno_Diet =  sapply(strsplit(data$Sample,"_"), function(x) x[2])
data$Geno = tolower(gsub("-.*","",data$Geno_Diet))
data$Diet = gsub(".*-","",data$Geno_Diet)
data$ID = gsub("_.*","",data$Sample)


data$ID = paste0("S",as.numeric(substr(data$ID,2,15)))

data = left_join(data,meta,by="ID")


table(data$Geno,data$Diet,data$Sex)

data = subset(data, select= - c(Geno_Diet))



```



### Missingness

Here we create a heatmap of missingness to see if it specific to any group of interest (Diet, Genotype). If there is any specificity, you would see very clear clustering blocks. If nothing shows up, there were no missing values


```{r, fig.height = 9,fig.width=9}
#Replace 0s with NA
data[data==0]=NA

miss.df = data
rownames(miss.df) = miss.df$Sample
miss.df = miss.df[,colnames(miss.df) %in% Molecule]
miss.df[!is.na(miss.df)] = 1
miss.df[is.na(miss.df)] = 0

missing = apply(miss.df,2, function(x) sum(x)/length(x))
anno = data.frame(Genotype = data$Geno,Diet = data$Diet,Sex=data$Sex)
rownames(anno) = rownames(miss.df)

if( length(which(missing<0.95)) > 0){
  pheatmap(t(miss.df[,which(missing<0.95)
  ]),color=c("white","black"),annotation_col = anno)
}

```


### Testing for specificity

There are not enough samples to really startify by sex and test for specificty, so this analysis only adjusts for sex

```{r, include=F}

miss.df$Geno = data$Geno
miss.df$Diet = data$Diet
miss.df$Sex = data$Sex
colnames(miss.df) = make.names(colnames(miss.df))
res.diet=NULL
res.geno=NULL
res.delta.delta=NULL

for(i in colnames(miss.df[,which(missing<0.95)])){
  
  f=as.formula(paste0(i,"~ Diet*Geno"))
  fit = glm(f, data = miss.df, family = "binomial", method = brglmFit)
  
  res.diet = rbind(res.diet,data.frame(Compound = i,summary(emmeans(fit,pairwise~Diet|Geno,adjust="none")$contrasts)))
  
  res.geno = rbind(res.geno,data.frame(Compound = i,summary(emmeans(fit,pairwise~Geno|Diet,adjust="none")$contrasts)))
  
  res.delta.delta = rbind(res.delta.delta,data.frame(Compound = i,joint_tests(fit)[3,]))
  res.delta.delta$model.term = "Delta - Delta"

}


res.geno$FDR = p.adjust(res.geno$p.value,method="BH")

res.diet$FDR = p.adjust(res.diet$p.value,method="BH")

res.delta.delta$FDR = p.adjust(res.delta.delta$p.value,method="BH")



```

#### If rows are empty or nothing is showing, no specificty was detected

```{r}

if(length(colnames(miss.df[,which(missing<0.95)]))>0){

  knitr::kable(res.diet[res.diet$FDR < 0.1,],caption="Compounds specific to a Diet")
  knitr::kable(res.geno[res.geno$FDR < 0.1,],caption="Compounds specific to a Genotype")
  knitr::kable(res.delta.delta[res.delta.delta$FDR < 0.1,],caption="Compounds specific to a Diet/Genotype Combo")
}


```

### Removed for too much missgingness

```{r}

#These compounds are being removed for excessive missingness
knitr::kable(names(missing[missing<0.3]))

#drop heavily missing data (>30%)
data = data[, ! colnames(data) %in% names(missing[missing<0.3])]

```


```{r}

#impute
data[,colnames(data) %in% Molecule] = t(imputeLCMD::impute.QRILC(log2(t(data[,colnames(data) %in% Molecule])))[[1]])

```

## Exploratory {.tabset}
### PCA

```{r, fig.height=6,fig.width=16}

pc.plot = function(df){
  pc = prcomp(data.matrix(df[,colnames(df) %in% Molecule]))
  pc.df = as.data.frame(pc$x)
  pc.df$Genotype = df$Geno
  pc.df$Diet = df$Diet
  
  p1 = ggplot(pc.df,aes(x=PC1,y=PC2,color = Diet,shape=Genotype, linetype=Genotype )) +
    geom_point() +
    geom_mark_ellipse() +
    theme_bw(12) +
    scale_color_nejm()

  p2 = ggplot(pc.df,aes(x=PC3,y=PC4,color = Diet,shape=Genotype, linetype=Genotype )) +
    geom_point() +
    geom_mark_ellipse() +
    theme_bw(12) +
    scale_color_nejm()

  p3 = ggplot(pc.df,aes(x=PC5,y=PC6,color = Diet,shape=Genotype, linetype=Genotype )) +
    geom_point() +
    geom_mark_ellipse() +
    theme_bw(12) +
    scale_color_nejm()

  print(p1+p2+p3)

}

pc.pool = pc.plot(data) +  plot_annotation('Pooled Sexes',theme=theme(plot.title=element_text(hjust=0.5)))

ggsave(pc.pool,file=paste0(gsub("\\..*","", file),"/pooled_pc.tif"))

pc.m = pc.plot(data[toupper(data$Sex) == "M",]) +  plot_annotation('Males Only',theme=theme(plot.title=element_text(hjust=0.5)))

ggsave(pc.m,file=paste0(gsub("\\..*","", file),"/male_pc.tif"))

pc.f = pc.plot(data[toupper(data$Sex) == "F",]) +  plot_annotation('Females Only',theme=theme(plot.title=element_text(hjust=0.5)))

ggsave(pc.f,file=paste0(gsub("\\..*","", file),"/female_pc.tif"))

```



### Heatmap {.tabset}
#### Pooled

```{r, fig.width=9,fig.height=25}

anno = data.frame(Genotype = data$Geno, data$Diet,Sex = data$Sex)
rownames(anno) = data$Sample
rownames(data) = data$Sample
pheatmap(t(log2(data[,colnames(data) %in% Molecule])),
         annotation_col = anno,
         scale="row",
         color=colorRampPalette(c("blue","white","red"))(100))


```


#### Males Only

```{r, fig.width=9,fig.height=25}

anno = data.frame(Genotype = data[data$Sex == "M",]$Geno,Diet = data[data$Sex == "M",]$Diet)
rownames(anno) = data[data$Sex == "M",]$Sample

pheatmap(t(log2(data[data$Sex == "M",colnames(data) %in% Molecule])),
         annotation_col = anno,
         scale="row",
         color=colorRampPalette(c("blue","white","red"))(100))



```



#### Females Only

```{r, fig.width=9,fig.height=25}

anno = data.frame(Genotype = data[data$Sex == "F",]$Geno,Diet = data[data$Sex == "F",]$Diet)
rownames(anno) = data[data$Sex == "F",]$Sample

pheatmap(t(log2(data[data$Sex == "F",colnames(data) %in% Molecule])),
         annotation_col = anno,
         scale="row",
         color=colorRampPalette(c("blue","white","red"))(100))



```


## Differntial Abundance {.tabset}

### Pooled {.tabset}

```{r, fig.height=14,fig.width=14}


data$Geno_Diet = paste0(data$Geno,data$Diet)

design = model.matrix(~ 0 + Geno_Diet + Sex, data= data)
fit <- lmFit(t(data[,colnames(data) %in% Molecule]), design, robust = TRUE)
cont2 = makeContrasts( Geno_DietkiHFD - Geno_DietwtHFD ,
                       Geno_DietkiLFD- Geno_DietwtLFD,
                       (Geno_DietkiHFD - Geno_DietwtHFD) -
                         (Geno_DietkiLFD- Geno_DietwtLFD),
                       Geno_DietkiHFD - Geno_DietkiLFD,
                       Geno_DietwtHFD - Geno_DietwtLFD,
                       levels = design)
set.seed(777)
fit.cont <- contrasts.fit(fit, contrasts = cont2) 
set.seed(777)
ebayesf <- eBayes(fit.cont, robust = T, trend = T)
tops.sig = topTable(ebayesf, adjust.method = "BH", p.value = 0.05, number = Inf)
tops.all = topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf)
```

#### HFD KI - HFD WT

```{r, fig.height=9,fig.width=9}

hfd.df = data.frame(Contrast = "KI HFD vs WT HFD", topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf,coef=1))

plot_table_save(hfd.df)

```


#### LFD KI - HFD WT

```{r, fig.height=9,fig.width=9}

lfd.df = data.frame(Contrast = "KI lfd vs WT lfd", topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf,coef=2))
plot_table_save(lfd.df)


```



####  [KI HFD vs. LFD] vs. [WT HFD vs. LFD]

```{r, fig.height=9,fig.width=9}

delta.delta.hfd = data.frame(Contrast = "Delta-Delta", topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf,coef=3))

plot_table_save(delta.delta.hfd)


```


####  KI HFD - KI LFD

```{r, fig.height=9,fig.width=9}

kihl.df = data.frame(Contrast = "KI hfd vs KI lfd", topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf,coef=4))

plot_table_save(kihl.df)


```


####  WT HFD - WT LFD

```{r, fig.height=9,fig.width=9}

wthl.df = data.frame(Contrast = "WT hfd vs WT lfd", topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf,coef=5))

plot_table_save(wthl.df)


```



#### Boxplots Top Compounds

```{r, fig.height=12,fig.width=16}

 
plot_differences(data,"Pooled")


```


### Males {.tabset}

```{r, fig.height=14,fig.width=14}

design = model.matrix(~ 0 + Geno_Diet , data= data[data$Sex == "M", ])


fit <- lmFit(t(data[data$Sex == "M",colnames(data) %in% Molecule]), design, robust = TRUE)

cont2 = makeContrasts( Geno_DietkiHFD - Geno_DietwtHFD ,
                       Geno_DietkiLFD- Geno_DietwtLFD,
                       (Geno_DietkiHFD - Geno_DietwtHFD) -
                         (Geno_DietkiLFD- Geno_DietwtLFD),
                       Geno_DietkiHFD - Geno_DietkiLFD,
                       Geno_DietwtHFD - Geno_DietwtLFD,
                       levels = design)
set.seed(777)
fit.cont <- contrasts.fit(fit, contrasts = cont2) 

set.seed(777)
ebayesf <- eBayes(fit.cont, robust = T, trend = T)

tops.sig = topTable(ebayesf, adjust.method = "BH", p.value = 0.05, number = Inf)
tops.all = topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf)
```

#### HFD KI - HFD WT

```{r, fig.height=9,fig.width=9}

hfd.df.m = data.frame(Contrast = "Male KI HFD vs WT HFD", topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf,coef=1))

plot_table_save(hfd.df.m)


```


#### LFD KI - LFD WT

```{r, fig.height=9,fig.width=9}

lfd.df.m = data.frame(Contrast = "Male KI lfd vs WT lfd", topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf,coef=2))

plot_table_save(lfd.df.m)


```



####  [KI HFD vs. LFD] vs. [WT HFD vs. LFD]

```{r, fig.height=9,fig.width=9}

detla.delta.df.m = data.frame(Contrast = "Male Delta-Delta", topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf,coef=3))

plot_table_save(detla.delta.df.m)



```


####  KI HFD - KI LFD

```{r, fig.height=9,fig.width=9}

kihl.df.m = data.frame(Contrast = "Male KI hfd vs KI lfd", topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf,coef=4))

plot_table_save(kihl.df.m)

```


####  WT HFD - WT LFD

```{r, fig.height=9,fig.width=9}

wthl.df.m = data.frame(Contrast = "Male WT hfd vs WT lfd", topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf,coef=5))

plot_table_save(wthl.df.m)

```



#### Boxplots Top Compounds

```{r, fig.height=12,fig.width=16}

plot_differences(data[data$Sex == "M",],level="Male") 
```

### Females {.tabset}

```{r, fig.height=14,fig.width=14}

design = model.matrix(~ 0 + Geno_Diet , data= data[data$Sex == "F", ])

fit <- lmFit(t(data[data$Sex == "F",colnames(data) %in% Molecule]), design, robust = TRUE)

cont2 = makeContrasts( Geno_DietkiHFD - Geno_DietwtHFD ,
                       Geno_DietkiLFD- Geno_DietwtLFD,
                       (Geno_DietkiHFD - Geno_DietwtHFD) -
                         (Geno_DietkiLFD- Geno_DietwtLFD),
                       Geno_DietkiHFD - Geno_DietkiLFD,
                       Geno_DietwtHFD - Geno_DietwtLFD,
                       levels = design)

set.seed(777)
fit.cont <- contrasts.fit(fit, contrasts = cont2) 

set.seed(777)
ebayesf <- eBayes(fit.cont, robust = T, trend = T)

tops.sig = topTable(ebayesf, adjust.method = "BH", p.value = 0.05, number = Inf)
tops.all = topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf)
```

#### HFD KI - HFD WT

```{r, fig.height=9,fig.width=9}

hfd.df.f = data.frame(Contrast = "Female KI HFD vs WT HFD", topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf,coef=1))

plot_table_save(hfd.df.f)

```


#### LFD KI - LFD WT

```{r, fig.height=9,fig.width=9}

lfd.df.f = data.frame(Contrast = "Female KI lfd vs WT lfd", topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf,coef=2))

plot_table_save(lfd.df.f)

```



####  [KI HFD vs. LFD] vs. [WT HFD vs. LFD]

```{r, fig.height=9,fig.width=9}

delta.delta.df.f = data.frame(Contrast = "Female Delta-Delta", topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf,coef=3))

plot_table_save(delta.delta.df.f)

```

####  KI HFD - KI LFD

```{r, fig.height=9,fig.width=9}

kihl.df.f = data.frame(Contrast = "Female KI hfd vs KI lfd", topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf,coef=4))

plot_table_save(kihl.df.f)

```


####  WT HFD - WT LFD

```{r, fig.height=9,fig.width=9}

wthl.df.f = data.frame(Contrast = "Female WT hfd vs WT lfd", topTable(ebayesf, adjust.method = "BH", p.value = 1, number = Inf,coef=5))

plot_table_save(wthl.df.f)

```


#### Boxplots Top Compounds

```{r, fig.height=12,fig.width=16}

plot_differences(data[data$Sex == "F",],level="Female") 


```
